--- 
title: "Writing and Rewriting for Scientific Communication"
subtitle: "A Workshop for the Anderson Research Group"
author: "Brooke Anderson"
date: "March 28, 2019"
site: bookdown::bookdown_site
output:
  bookdown::tufte_html_book:
    toc: yes
    css: toc.css
    tufte_variant: "envisioned"
  bookdown::tufte_book2:
    toc: yes
    includes:
      in_header: preamble.tex
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "Online booklet for one-week writing workshop for students in my research group."
---

```{r include = FALSE}
library(tufte)
library(knitr)
```

# Introduction

`r newthought("Welcome to our one-week")` writing workshop! This workshop is
meant for students who are working on a manuscript that they plan to submit for
peer review at a scientific journal. We'll be focusing on how to structure and
rewrite your manuscript to make it clearer and easier to read. 

There are a number of things that are important that we won't cover here: 

1. **Grammar and usage.** These are both very important, and you'll want these
to be impeccable when you submit your manuscript. However, there are loads of
resources and courses for this, and while I often see cases where students 
haven't mastered this yet, I don't think it's hard to find the resources to 
learn these. Also, at least through your graduate career, you will almost always
have co-authors, including a senior author (typically the head of your research 
group). Grammar and usage problems are easy to fix quickly when your senior author
edits your paper. The things we will cover here take more time and work. 
2. **Doing a literature review.** This is an art of its own. It's very important
for your manuscript to use good examples from the literature to set up why your
study is important and to put what you find in the context of previously
published studies. If you are pursuing a graduate degree, you will need to
explore the literature enough to understand the current state of science in your
field and how your work fits into the field. With certain types of manuscripts,
like commentaries and reviews, the whole manuscript might be a synthesis and
exploration of the literature. Ideally, you will have done quite a bit of 
work to explore and get examples from the literature by the time we do this 
workshop, as this will provide some of the raw material for your manuscript. 
However, we will only tangentially discuss how to explore and review the literature. 
This will be in the context of building and presenting arguments in your manuscript, 
as evidence from the literature will often be needed to do this, particularly in the
Introduction and Discussion.
3. **Creating effective figures and tables.** This is also extremely important. 
At worst, bad figures and tables can make a manuscript look unprofessional. If
they do, then a reviewer may worry that the research presented in the study is 
similarly sloppy or poorly done. Figures and tables are the best chance to immediately
engage a reader, and they'll often stick with the reader, if they're good, much 
longer than the text. They also served double-duty, as you'll often also include
them in presentations you make about the research. Finally, a really good figure or
table encourages the reader to explore the results from the research deeply, and 
start looking for patterns or trends that suggest ideas for next steps for the 
research area. Manuscripts are more likely to get cited, and to have a higher 
impact on your field, if readers leave the paper with ideas of research they want
to do to take the next steps. There are some excellent resources for learning how
to design and create figures and tables. I particularly like books by Edward Tufte
and Howard Wainer. There are also a number of books on visualization that are 
specific to a programming language, like R or Python.
4. **How to get words on paper.** Our focus will be strongly on improving and
revising a manuscript once you've already got some stuff down on paper. We won't
talk much about how to get some stuff down in the first place. There are some
good books on the topic of writing initial drafts. Different writers have
different styles for this. Some will work very hard on the first draft, so that
rewriting doesn't need to be extensive. I tend to start by writing a lot, but
all of it pretty messy and drafty, and then wrangling and editing that raw
material into a more solid draft. I usually do lots of rounds of drafts,
starting with editing to move big parts around and form the larger structure of
the paper and to see where there are holes that need more evidence or support.
Then, in later drafts, my editing process involves drilling down to make
the writing clearer and easier to read. In this workshop we'll focus on the work
involved with the later stages of editing. If you'd like to learn more about
tips and techniques for earlier stages, where the focus is generating your
starting material, I like the books of Peter Elbow. Stephen King's *On Writing*
is also great, particularly in inspiring you to get something down even when 
you're discouraged (by bad reviews or by how big the task seems, for example).

Click on the **Next** button (or navigate using the
links at the top of the page) to continue.

## Readings for the workshop

We will be reading and trying out the ideas from several books for this
workshop. They all have connections to Joseph Williams *Style* book, which has
gone through several editions (and titles) since it was first published in the
1980s. 

There are a lot of good writing books out there, but this is by far my favorite. 
I regularly pick it back up when I am editing manuscripts, particularly when I have
a section of a draft that I am struggling to write clearly. What I like about it
is that it focuses on how we can edit our writing to make it easier for readers
to read and understand. It bases its advice on how readers process information as
they read, and it gives advice that is easy to use to diagnose problems and
fix them. 

I cannot overstate how important it is to edit your scientific manuscript until
it can be read easily (and feel "quick" to read) and until it is hard to
misunderstand. You are moving into work where you will be the expert in the
room. You will know your very specific topic better than all but a handful of
people in the world, and often better than the editor and reviewers 
for your manuscript. You are teaching through your manuscript, and the text 
must be clear for readers to learn what you've learned. 

Also, scientific papers are hard work to read, even when they are written
beautifully. As the reader, you must work to understand the science and
statistical methods that were used to generate results for the paper. Your mind
works hard to integrate the paper's results with what you know from other papers
in the field. If the writer does not make the writing clear and easy to read, it
can feel impossible to read the paper, because your brain can't simultaneously
try to figure out what the writer was trying to say and what the science in the
paper is saying. You have worked hard on your science. It's worth the effort to
edit your manuscript until you remove all overhead you can that makes the
writing hard to read. Most people, even senior scientists, write first drafts
that are a drag to read. Even if all the information is there, it's a lot of
work to mentally extract it, and it's often hard to follow along with the
arguments and explanations. It's fine for your early drafts to be like that.
It's not good for the submitted manuscript to still feel so heavy.

The examples in *Style* are mainly from academic non-fiction, although not 
specifically from scientific manuscripts or manuscripts meant for peer review. 
Fortunately, two other writers have taken Williams's ideas and illustrated 
how they work in scientific writing. Further, they give some tips specific to 
scientific writing, including how to plan and structure certain sections---like the
Introduction and Discussion. 

The final book includes a section on arguments. We often don't think of
scientific writing as argumentative, but it really is. You will need to convince
readers that your study is important, that it advances the science. Then you'll
need to not just present your results, but provide an interpretation for those
results. You will need to walk your readers through why those interpretations
are reasonable and appropriate. While many students come into graduate school
with a reasonable mastery of grammar, they often present their arguments in a
way that is scattered and diffuse. Often, when I am editing manuscripts with
first authors who are students, much of the editing process is related to
finding and fleshing out their arguments. We won't be able to go deeply into 
rhetoric in this workshop, but we will cover the basics of building arguments
in your writing based on the advice in this book.

The full list of books for the workshop is: 

1. [*Style: The Basics of Clarity and Grace (2nd edition)* by Joseph
Williams.](https://www.amazon.com/Style-Basics-Clarity-Grace-2nd/dp/0321330854/ref=sr_1_20?dchild=1&keywords=joseph+williams+style&qid=1594599265&s=books&sr=1-20)
(This edition is no longer in print. I prefer the earlier editions, which are
shorter. There tend to always be used copies available on Amazon. There also
seems to be a pdf of an early edition version [available
online](https://sites.duke.edu/niou/files/2014/07/WilliamsJosephM1990StyleTowardClarityandGrace.pdf)
that you can use before your print edition arrives.)
2. [*Writing Science in Plain English* by Anne E.
Greene.](https://www.amazon.com/Writing-Science-English-Chicago-Publishing/dp/022602637X/ref=sr_1_2?crid=3R5FVEBHLEE6M&dchild=1&keywords=writing+science+in+plain+english&qid=1594599082&sprefix=writing+science+in+p%2Caps%2C185&sr=8-2)
3. [*Writing Science* by Joshua
Schimel](https://www.amazon.com/Writing-Science-Papers-Proposals-Funded/dp/0199760241/ref=pd_sbs_14_2/140-7397953-6612456?_encoding=UTF8&pd_rd_i=0199760241&pd_rd_r=a18d3605-570b-4f9c-a9b7-560aab4c2f19&pd_rd_w=reAtV&pd_rd_wg=ldykT&pf_rd_p=bdc67ba8-ab69-42ee-b8d8-8f5336b36a83&pf_rd_r=6JENX0H17SA0SEH0B3SH&psc=1&refRID=6JENX0H17SA0SEH0B3SH)
4. [*The Craft of Research (4th edition)* by Wayne C. Booth, Gregory C.
Colomb, Joseph M. Williams, Joseph Bizup, and William T.
FitzGerald.](https://www.amazon.com/Research-Chicago-Writing-Editing-Publishing/dp/022623973X/ref=sr_1_2?dchild=1&keywords=craft+of+research&qid=1594599147&s=books&sr=1-2)

We will also use three published articles as examples. None of these are traditional research
articles, but rather cover a range of commentary-/editorial-style articles. All three should be 
available either directly online or through CSU. Please let me know if you have any problem
accessing any of these articles.

1. [Gall, Melanie, Kevin A. Borden, and Susan L. Cutter. "When do losses count?
Six fallacies of natural hazards loss data." Bulletin of the American
Meteorological Society 90.6 (2009):
799-810.](https://journals.ametsoc.org/bams/article/90/6/799/59608/When-Do-Losses-Count-Six-Fallacies-of-Natural)
2. [Cheplygina, Veronika, et al. "Ten simple rules for getting started on
Twitter as a scientist." PLoS Computational Biology 16.2 (2020):
e1007513-e1007513.](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1007513)
3. [Schwartz, Sharon. "The fallacy of the ecological fallacy: the potential
misuse of a concept and the consequences." American Journal of Public Health
84.5 (1994):
819-824.](https://ajph.aphapublications.org/doi/abs/10.2105/AJPH.84.5.819)

I have picked these papers because they all provide important ideas and advice, and
they all have some strong writing. In particular, they have a clear overarching structure
and state and defend their arguments well. Also, like most published papers (mine included), 
they include examples where sentences or paragraphs could be edited to be easier to read
and harder to misunderstand. They therefore provide some nice examples of both the good and
the bad (none get into "the ugly", but you won't have to look too far in the scientific 
literature to find some that do), and so we can see that even very good papers have room
for further editing and try our hand at doing that. 

## Workshop schedule

Each day of the week, we will focus on a different topic.

Schedule: 




```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'ggplot2', 'readr',
  'forcats', 'magrittr', 'dplyr', 'sf', 'htmlwidgets', 'DT', 'plotly', 
  'leaflet', 'flexdashboard'
), 'packages.bib')
```

